---
title: "HW4"
date: '2023-04-24'
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    keep_md: yes
  pdf_document:
    toc: true
  always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE)
options(digits = 4)
```

```{r packages}
# Load packages
pacman::p_load(tidyverse, ivreg, ivmodel, magrittr, ggplot2, MASS)

```


### MonteCarlo Function

```{r parameters}
# Set parameters
b = 1
se = 1
sv = 1
p = .99
l = 4
n = 100 
mu_z = rep(0, l)
Qzz = diag(l)
N = 10000
```

```{r montecarlo, dependson="parameters"}
# Monte-Carlo function set-up
montecarlo = function(N, gamma){
  # Create empty vectors to put values into
  sim.beta = rep(NA, N)
  sim.se = rep(NA, N)
  t_stat = rep(NA, N)
  ci_lower = rep(NA, N)
  ci_upper = rep(NA, N)
  and_rub = rep(NA, N)

  # Montecarlo loop. This will output all the statistics the question asked for
  for (sim in 1:N){
    # Simulate z's
    z = mvrnorm(n, mu_z, Qzz)
    # Simulate errors
    e_v = mvrnorm(n, rep(0, 2), matrix(c(se^2, rep(p, 2), se^2), ncol = 2))
    
    # Create x's
    x = z %*% gamma + e_v[,2]
    # Create y's
    y = x %*% b + e_v[,1]
    
    # Calculate 2SLS
    tsls = ivreg(y ~ x | z)
    
    # Store values of interest from regression
    sim.beta[sim] = coef(tsls)[2] 
    sim.se[sim] = broom::tidy(tsls)[2,3] |> as.numeric()
    
    # Calculate t-stat
    t_stat[sim] = (sim.beta[sim] - b) / sim.se[sim]
    
    # Calculate confidence interval
    ci_lower[sim] = sim.beta[sim] - 1.96*sim.se[sim]
    ci_upper[sim] = sim.beta[sim] + 1.96*sim.se[sim]
    
    # Create z projection matrix for use in anderson-rubin stat
    pz = z %*% solve(t(z)%*%z)%*%t(z)
    
    # IVmodel object for anderson-rubin function
    iv = ivmodel(y, x, z)
    
    # Calculate anderson-rubin stat
    test = AR.test(iv, beta0 = 1)
    and_rub[sim] = test$Fstat
    

  }

  # Store sensibly
  tsls_output = tibble(beta = sim.beta, 
                         se = sim.se,
                         t_stat = t_stat,
                         ci_lower = ci_lower,
                         ci_uppwer = ci_upper,
                         anderson_rubin = and_rub)
  
  # Decision for t-tests. 1 for reject 0 for fail to reject. Using this convention because it is   convenient for calculating the percent of rejections
  tsls_output %<>% mutate(t_decision = if_else(abs(t_stat) > 1.96, 1, 0))
  
  # Same for AR tests with the same convention
  f_crit = qf(0.95, l, n-l)
  tsls_output %<>% mutate(ar_decision = if_else(anderson_rubin > f_crit, 1,0))
  
  return(tsls_output)

}
```

### Instrument Tests

#### Irrelevant Instruments

```{r irrelevant-instruments, dependson=c("parameters", "montecarlo")}
# Gamma is all zeros
gamma = rep(0, 4)

set.seed(123)
irrelevant = montecarlo(N, gamma)

# Histogram of 2sls with irrelevant instruments
irrelevant_b = irrelevant |> ggplot(aes(x = beta)) +
  geom_histogram(bins = 25) + 
  labs(x = "B 2SLS", y = "", title = "Distribution of Beta 2SLS with Irrelevant Instruments",
       subtitle = "Normal N(0,1) T-Distribution in Red",
       caption = "Gamma = (0,0,0,0)\', N = 10,000") +
  cowplot::theme_cowplot()

# Calculate the bias
bias_irrelevant = mean(irrelevant$beta) - b

# Histogram of t-stats
irrelevant_t = irrelevant |> ggplot(aes(x = t_stat)) + 
  geom_histogram(bins = 25) + 
  geom_histogram(data = data.frame(x = rnorm(N)), aes(x = x), fill = 'red', alpha = 0.5) + 
  labs(x = "T-Stat", y = "", title = "Distribution of T-statistics with Irrelevant Instruments",
       caption = "Gamma = (0,0,0,0)\', N = 10,000") +
  cowplot::theme_cowplot()

# Empirical size of t_tests
size_t_irr = sum(irrelevant$t_decision) / N

# Empirical size of Anderson-Rubin test
size_ar_irr = sum(irrelevant$ar_decision) / N
```

##### Printout of results
```{r dependson="irrelevant-instruments"}
head(irrelevant, 15)
```


##### Part 1: Histogram of beta hat 2SLS
```{r fig.align='left', dependson="irrelevant-instruments"}
irrelevant_b
```

##### Part 2: Bias 
For the irrelevant instruments case, the bias is equal to `r bias_irrelevant`.

##### Part 3: Histogram of T-Statistics
```{r fig.align='left', message=FALSE, warning=FALSE, dependson="irrelevant-instruments"}
irrelevant_t
```

##### Part 4: Empirical Size of T-Test
The empirical size of the t-test is `r size_t_irr`, while for the Anderson-Rubin test it is `r size_ar_irr`

#### Weak Instruments

```{r weak-instruments, dependson=c("parameters", "montecarlo")}
# Gamma is 0.1 and then zeros
gamma = c(0.1, rep(0, 3))
v_beta_hat = solve(t(gamma)%*%gamma)

set.seed(123)
weak = montecarlo(N, gamma)

# Histogram of 2sls with weak instruments
weak_b = weak |> ggplot(aes(x = beta)) +
  geom_histogram(bins = 25) + 
  geom_histogram(data = data.frame(x = rnorm(N, b, sqrt(v_beta_hat/N))),
                 aes(x = x), fill = 'red', alpha = 0.5) + 
  labs(x = "B 2SLS", y = "", title = "Distribution of 2SLS with Weak Instruments",
       subtitle = "Asymptotic Distribution in Red",
       caption = "Gamma = (0.1,0,0,0)\', N = 10,000") +
  cowplot::theme_cowplot()

# Calculate the bias
bias_weak = mean(weak$beta) - b

# Histogram of t-stats
weak_t = weak |> ggplot(aes(x = t_stat)) + 
  geom_histogram(bins = 25) + 
  geom_histogram(data = data.frame(x = rnorm(N)), aes(x = x), fill = 'red', alpha = 0.5) + 
  labs(x = "T-Stat", y = "", title = "Distribution of T-Statistics with Weak Instruments",
       subtitle = "Normal N(0,1) T-Distribution in Red",
       caption = "Gamma = (0.1,0,0,0)\', N = 10,000") +
  cowplot::theme_cowplot()

# Empirical size of t_tests
size_t_weak = sum(weak$t_decision) / N

# Empirical size of Anderson-Rubin test
size_ar_weak = sum(weak$ar_decision) / N
```

##### Printout of results
```{r dependson="weak-instruments"}
head(weak, 15)
```


##### Part 1: Histogram of beta hat 2SLS
```{r fig.align='left',dependson="weak-instruments"}
weak_b
```

##### Part 2: Bias 
For the weak instruments case, the bias is equal to `r bias_weak`.

##### Part 3: Histogram of T-Statistics
```{r fig.align='left', message=FALSE, warning=FALSE,dependson="weak-instruments"}
weak_t
```

##### Part 4: Empirical Size of T-Test
The empirical size of the t-test is `r size_t_weak`, while for the Anderson-Rubin test it is `r size_ar_weak`

#### Strong Instruments

```{r strong-instruments, dependson=c("parameters", "montecarlo")}
# Gamma is 1 then zeros
gamma = c(1, rep(0, 3))
v_beta_hat = solve(t(gamma)%*%gamma)

set.seed(123)
strong = montecarlo(N, gamma)

# Histogram of 2sls with strong instruments
strong_b = strong |> ggplot(aes(x = beta)) +
  geom_histogram(bins = 25) + 
  geom_histogram(data = data.frame(x = rnorm(N, b, sqrt(v_beta_hat/N))),
                 aes(x = x), fill = 'red', alpha = 0.5) + 
  labs(x = "B 2SLS", y = "", title = "Distribution of 2SLS with Strong Instruments",
       subtitle = "Asymptotic Distribution in Red",
       caption = "Gamma = (1,0,0,0)\', N = 10,000") +
  cowplot::theme_cowplot()

# Calculate the bias
bias_strong = mean(strong$beta) - b

# Histogram of t-stats
strong_t = strong |> ggplot(aes(x = t_stat)) + 
  geom_histogram(bins = 25) + 
  geom_histogram(data = data.frame(x = rnorm(N)), aes(x = x), fill = 'red', alpha = 0.5) + 
  labs(x = "T-Stat", y = "", title = "Distribution of T-Statistic with Strong Instruments",
       subtitle = "Normal N(0,1) T-Distribution in Red",
       caption = "Gamma = (1,0,0,0)\', N = 10,000") +
  cowplot::theme_cowplot()

# Empirical size of t_tests
size_t_strong = sum(strong$t_decision) / N

# Empirical size of Anderson-Rubin test
size_ar_strong = sum(strong$ar_decision) / N
```

##### Printout of results
```{r dependson="strong-instruments"}
head(strong, 15)
```


##### Part 1: Histogram of beta hat 2SLS
```{r fig.align='left',dependson="strong-instruments"}
strong_b
```

##### Part 2: Bias 
For the strong instruments case, the bias is equal to `r bias_strong`.

##### Part 3: Histogram of T-Statistics
```{r fig.align='left', message=FALSE, warning=FALSE,dependson="strong-instruments"}
strong_t
```

##### Part 4: Empirical Size of T-Test
The empirical size of the t-test is `r size_t_strong`, while for the Anderson-Rubin test it is `r size_ar_strong`


```{r}
ldkjaf;ds
```

















