---
title: "HW1"
output:
  pdf_document: default
  html_document: default
date: '2023-03-30'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      cache = TRUE)
```

```{r}
# Load packages
pacman::p_load(tidyverse, boot)
```

###  Question 1 Monte Carlo Simulation

```{r}
# Set true values of parameters
y_true = c(1, 1)
sigma = matrix(diag(c(0.25^2,1)), ncol = 2)
theta = y_true[1]/y_true[2]

# Simulating parameters
n = 50
B = 10000

# Set seed
set.seed(123)

# Initialize simulation vectors
sim.theta = rep(0, B)
sim.mu1 = rep(0,B)
sim.mu2 = rep(0, B)
sim.sigma = matrix(0, B, 3)

# Monte Carlo loop
for (sim in 1:B) {
  # Simulate data
  sim.y = tibble(y1 = rnorm(n, mean = y_true[1], sd = sigma[1,1]),
                 y2 = rnorm(n, mean = y_true[2], sd = sigma[2,2]))
  mu_hat = c(mean(sim.y$y1), mean(sim.y$y2))
  sigma_hat = with(sim.y, c(sd(y1), cov(y1, y2), sd(y2)))

  # Parameter of interest
  theta_hat = mu_hat[1]/mu_hat[2]
  
  # Store the simulated values 
  sim.theta[sim] = theta_hat
  sim.mu1[sim] = mu_hat[1]
  sim.mu2[sim] = mu_hat[2]
  sim.sigma[sim,] = sigma_hat
}

# Calculate the mean and standard error of theta_hat
mean_theta_hat = mean(sim.theta); mean_theta_hat
se_theta_hat = sd(sim.theta); se_theta_hat

# Confidence interval of theta_hat
mean_theta_hat + c(-se_theta_hat, se_theta_hat)*1.96

# Var cov matrix
sigma_hat = c(mean(sim.sigma[,1]), rep(mean(sim.sigma[,2]), 2), mean(sim.sigma[,3])) |> matrix(nrow = 2)
```

### Question 2

Theta is defined as $\theta = \frac{\mu_1}{\mu_2} = f(\mu)$. To use the delta method, we take the derivative of $\theta$ in therms of $\mu_1 \text{ and } \mu_2$.

$$
g(\mu) = \frac{\partial f(\mu)}{\partial\mu} = \left(\begin{array}{c}
\frac{\partial f(\mu)}{\partial \mu_1}  \\
\frac{\partial f(\mu)}{\partial \mu_2}  
\end{array}
\right) = \left(\begin{array}{c}
\frac{1}{\mu_2} \\
\frac{-\mu_1}{\mu_2^2}
\end{array}
\right)
$$

Using this, we can calculate the variance as

$$
\widehat{\mathrm{var}}(\hat{\theta}) = g(\hat{\theta})'\hat{V}_{\hat{\theta}}g(\hat{\theta})
$$

So the standard error is

$$
\widehat{\mathrm{se}}(\hat{\mu}) = \sqrt{\widehat{\mathrm{var}}(\hat{\theta})}
$$

We already calculated the $\hat{V}_{\hat{\theta}}$ in part 1, so we just need to enumerate $g(\theta)$

```{r}
g_mu_hat = c(1/y_true[1], -y_true[1]/y_true[2]) |> matrix()
varhat_theta_hat = t(g_mu_hat)%*%sigma_hat%*%g_mu_hat
sehat_theta_hat = sqrt(varhat_theta_hat) # This isn't right
```


### Question 3 Jackknife Method

```{r}
# Simulated data
y_sim = tibble(y1 = rnorm(n, mean = y_true[1], sd = sigma[1,1]), 
               y2 = rnorm(n, mean = y_true[2], sd = sigma[2,2]))

# Initialize values
n = nrow(y_sim)
jack_theta = rep(0, n)

# Set seed
set.seed(123)

# Jackknife loop
for (i in 1:n){
  # Selects all the data except the one observation we're leaving out
  temp_data = y_sim[-i,]
  
  # Calculate the parameters
  mu_hat = c(mean(temp_data$y1), mean(temp_data$y2))

  # Parameter of interest
  theta_hat = mu_hat[1]/mu_hat[2]

  # Store the parameter
  jack_theta[i] = theta_hat
}

# Jackknife standard error
jack_mean = mean(jack_theta)
jack_se = ((n-1)/n)*sum((jack_theta - theta)^2);jack_se 

# Confidence interval
jack_mean + c(-jack_se, jack_se)*1.96 

```

### Question 4 NonParametric Bootstrap

```{r}
# Bootstrap parameters
n = 50
B = 10000

# Initialize necessary vectors
boot_theta = rep(0, B)
list_boot_theta = vector('list', 5) # empty list to put results from different seeds into
boot_test = vector('list', 5)
# Bootstrap loop
# Outside loop is to loop over the different seeds
for (seed in 123:127){
  set.seed(seed)
  y_sim = tibble(y1 = rnorm(n, mean = y_true[1], sd = sigma[1,1]), 
                 y2 = rnorm(n, mean = y_true[2], sd = sigma[2,2]))
  # Inside loop is bootstrapping
  for (i in 1:B){
    # Get bootstrap sample
    idx = sample(n, replace = TRUE)
    boot_samp = y_sim[idx,]
    
    # Get parameters
    mu_hat = c(mean(boot_samp$y1), mean(boot_samp$y2))
    
    # Parameter of interest
    boot_theta[i] = mu_hat[1]/mu_hat[2]
    boot_test[[i]] = boot_samp
  }
  list_boot_theta[[seed-122]] = boot_theta # Subtract 122 so the list indexes from 1
}

# Parameter of interest for each seed
boot_theta_hat = sapply(1:5, function(i) {mean(list_boot_theta[[i]])})
boot_se_theta_hat = sapply(1:5, function(i) {sd(list_boot_theta[[i]])});boot_se_theta_hat

# Normal CI
sapply(1:5, function(i){boot_theta_hat[[i]] + c(-boot_se_theta_hat[[i]], boot_se_theta_hat[[i]])*1.96}) # CIs are read top to bottom. Going left to right are the five samples from 5 different seeds
```

To calculate the other CIs, we will use the bootstrap package

```{r}
# This is a function to get the parameter given a bootstrapped sample
fun_theta = function(x, i){
  temp_data = x[i,]
  boot_theta = mean(temp_data$y1)/mean(temp_data$y2)
  return(boot_theta)
}

# Initialize empty list for the result from the five seeds
list_boot_theta_ci = vector('list', 5)

# Loop over the seeds again
for (seed in 123:127) {
  set.seed(seed)
  y_sim = tibble(y1 = rnorm(n, mean = y_true[1], sd = sigma[1,1]), 
                 y2 = rnorm(n, mean = y_true[2], sd = sigma[2,2]))
  theta.boot = boot(y_sim, statistic=fun_theta,R=10000)
  
  list_boot_theta_ci[[seed-122]] = theta.boot
}

# All the confidence intervals
# Im not doing this in a loop because it takes away the nice formatting of the CIs
boot.ci(list_boot_theta_ci[[1]], conf=0.95, type=c("norm", "perc", "bca"))
boot.ci(list_boot_theta_ci[[2]], conf=0.95, type=c("norm", "perc", "bca"))
boot.ci(list_boot_theta_ci[[3]], conf=0.95, type=c("norm", "perc", "bca"))
boot.ci(list_boot_theta_ci[[4]], conf=0.95, type=c("norm", "perc", "bca"))
boot.ci(list_boot_theta_ci[[5]], conf=0.95, type=c("norm", "perc", "bca"))

# The normal CI is not quite the same as my manual one, but close enough I think

```

### Question 5 Parametric Bootstrap

```{r}

```











